### 2018-08-10 paper/acknowledgments.tex
paper from overleaf

% This is here for compliance with the DESC Publication Policy, as the paper will ultimately have to be compiled with the start_paper format.
%
%This is the text imported from \code{acknowledgments.tex}, and will be replaced by some standard LSST DESC boilerplate at some point.
%

The DESC acknowledges ongoing support from the Institut National de Physique Nucl\'eaire et de Physique des Particules in France; the Science \& Technology Facilities Council in the United Kingdom; and the Department of Energy, the National Science Foundation, and the LSST Corporation in the United States.
DESC uses resources of the IN2P3 Computing Center (CC-IN2P3--Lyon/Villeurbanne - France) funded by the Centre National de la Recherche Scientifique; the National Energy Research Scientific Computing Center, a DOE Office of Science User Facility supported by the Office of Science of the U.S.\ Department of Energy under Contract No.\ DE-AC02-05CH11231; STFC DiRAC HPC Facilities, funded by UK BIS National E-infrastructure capital grants; and the UK particle physics grid, supported by the GridPP Collaboration.
This work was performed in part under DOE Contract DE-AC02-76SF00515.

% hack event acknowledgments are important
We thank Stony Brook University for hosting the Summer 2017 LSST-DESC Hack Day at which this work was partially completed.

% individual acknowledgments for everyone's grants, probably should be in alphabetical order
AIM is advised by David Hogg and was supported by National Science Foundation grant AST-1517237 and the U.S. Department of Energy, Office of Science, Office of Workforce Development for Teachers and Scientists, Office of Science Graduate Student Research (SCGSR) program, administered by the Oak Ridge Institute for Science and Education for the DOE under contract number DE‐SC0014664.

JYHS is advised by Ofer Lahav and acknowledges support from the MyBrainSc Scholarship, provided by the Ministry of Education, Malaysia. JS would also like to thank Antonella Palmese and Boris Leistedt for guidance on the use of the algorithms \textsc{ANNz2} and \textsc{Delight} respectively.