### 2019-03-15 paper/main.tex
code macros

\RequirePackage[switch, columnwise, running, mathlines, displaymath,
mathlines]{lineno}
\RequirePackage{docswitch}
% \flag is set by the user, through the makefile:
%    make note
%    make apj
% etc.
\setjournal{\flag}

\documentclass[\docopts]{\docclass}

\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
  \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
  \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
  \renewenvironment{#1}%
     {\linenomath\csname old#1\endcsname}%
     {\csname oldend#1\endcsname\endlinenomath}}%
\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
  \patchAmsMathEnvironmentForLineno{#1}%
  \patchAmsMathEnvironmentForLineno{#1*}}%
\AtBeginDocument{%
\patchBothAmsMathEnvironmentsForLineno{equation}%
\patchBothAmsMathEnvironmentsForLineno{align}%
\patchBothAmsMathEnvironmentsForLineno{flalign}%
\patchBothAmsMathEnvironmentsForLineno{alignat}%
\patchBothAmsMathEnvironmentsForLineno{gather}%
\patchBothAmsMathEnvironmentsForLineno{multline}%
}

% You could also define the document class directly
%\documentclass[]{emulateapj}

% Custom commands from LSST DESC, see texmf/styles/lsstdesc_macros.sty
\usepackage{lsstdesc_macros}
% \usepackage{natbib}
% \usepackage{subcaption}
\usepackage[colorinlistoftodos]{todonotes}
% \usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{multirow}
\usepackage{xspace}
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage{wasysym}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{footmisc}
\graphicspath{{./}{./figures/}}
\bibliographystyle{mnras}%mn2e

% Add your own macros here:
\newcommand{\textul}{\underline}

% projects
\newcommand{\proj}[1]{\textsc{#1}}
\newcommand{\lsst}{\proj{LSST}}
\newcommand{\lsstdesc}{\lsst-\proj{DESC}}
\newcommand{\sdss}{\proj{SDSS}}
\newcommand{\buzz}{\proj{Buzzard}}

% can easily change these to specify posterior throughout, etc
\newcommand{\pz}{photo-$z$}
\newcommand{\pzpdf}{\pz\ PDF}
\newcommand{\Pz}{Photo-$z$}
\newcommand{\Pzpdf}{\Pz\ PDF}
\newcommand{\chisq}{$\chi^{2}$}

% codes
\newcommand{\pzcode}[1]{\texttt{#1}}
\newcommand{\qp}{\pzcode{qp}}
\newcommand{\annz}{\pzcode{ANNz2}}
\newcommand{\bpz}{\pzcode{BPz}}
\newcommand{\cmnn}{\pzcode{CMNN}}
\newcommand{\delight}{\pzcode{Delight}}
\newcommand{\eazy}{\pzcode{EAZY}}
\newcommand{\flexzboost}{\pzcode{FlexZBoost}}
\newcommand{\gpz}{\pzcode{GPz}}
\newcommand{\lephare}{\pzcode{LePhare}}
\newcommand{\metaphor}{\pzcode{METAPhoR}}
\newcommand{\skynet}{\pzcode{SkyNet}}
\newcommand{\tpz}{\pzcode{TPZ}}
\newcommand{\trainz}{\pzcode{trainZ}}

\newcommand*\mathinhead[2]{\texorpdfstring{$\boldsymbol{#1}$}{#2}}

\newcommand{\red}[1]{\textcolor{red}{#1}}
% Claim a color for your comments
\newcommand{\aim}[1]{\textcolor{green}{#1}}%Alex Malz comments in green
\newcommand{\blue}[1]{\textcolor{blue}{#1}}%unknown person comments in blue
\definecolor{scc}{rgb}{0.0, 0.26, 0.15}
\newcommand{\scc}[1]{\textcolor{scc}{#1}}%Stefano Cavuoti Comments in dark green
\definecolor{newpink}{rgb}{0.858, 0.188, 0.478}
\newcommand{\erfan}[1]{\textcolor{newpink}{#1}} % Erfan Nourbakhsh comments in pink
\newcommand{\jan}[1]{\textcolor{orange}{#1}}%Jeff Newman comments in Orange

\def\X{{\mathbf{X}}}
\def\x{{\mathbf{x}}}
\def\E{{\mathbb{E}}}

% ======================================================================

\begin{document}
\linenumbers

\title{Evaluation of probabilistic photometric redshift estimation approaches for \textsc{LSST}}

\maketitlepre

\begin{abstract}

Many scientific investigations of photometric galaxy surveys require redshift estimates, whose uncertainty properties are best encapsulated by photometric redshift (photo-$z$) posterior probability distribution functions (PDFs).
A plethora of photo-$z$ PDF estimation methodologies abound, producing discrepant results with no consensus on a preferred approach.
We present the results of a comprehensive experiment comparing twelve photo-$z$ algorithms on mock data produced for the Large Synoptic Survey Telescope (\textsc{LSST}) Dark Energy Science Collaboration (\textsc{DESC}).
By supplying perfect prior information, in the form of the complete template library and a representative training set as inputs to each code, we demonstrate the impact of the assumptions underlying each technique on the output photo-$z$ PDFs.
In the absence of a notion of true, unbiased photo-$z$ PDFs, we evaluate and interpret multiple metrics of the ensemble properties of the derived photo-$z$ PDFs as well as traditional reductions to photo-$z$ point estimates.
We report systematic biases and overall over/under-breadth of the photo-$z$ PDFs of many popular codes, which may indicate avenues for improvement in the algorithms or implementations.
Furthermore, we raise attention to the limitations of established metrics for assessing photo-$z$ PDF accuracy; though we identify the conditional density estimate (CDE) loss as a promising metric of photo-$z$ PDF performance in the case where true redshifts are available but true photo-$z$ PDFs are not, we emphasize the need for science-specific performance metrics.

\end{abstract}

% Keywords are ignored in the LSST DESC Note style:
\dockeys{galaxies: distances and redshifts -- galaxies: statistics -- methods: statistical}

\maketitlepost

% ----------------------------------------------------------------------
% {% raw %}

\input{./tex/introduction}

\input{./tex/data}

\input{./tex/methods}

\input{./tex/analysis}

\input{./tex/results}

\input{./tex/discussion}

\input{./tex/conclusion}

% ----------------------------------------------------------------------

\subsection*{Acknowledgments}

%%% Here is where you should add your specific acknowledgments, remembering that some standard thanks will be added via the \code{desc-tex/ack/*.tex} and \code{contributions.tex} files.

%This paper has undergone internal review in the LSST Dark Energy Science Collaboration. % REQUIRED if true

\input{contributions} % Standard papers only: author contribution statements. For examples, see http://blogs.nature.com/nautilus/2007/11/post_12.html

% This work used TBD kindly provided by Not-A-DESC Member and benefitted from comments by Another Non-DESC person.

% Standard papers only: A.B.C. acknowledges support from grant 1234 from ...

The authors would like to thank their LSST-DESC publication review committee for comments that improved the paper draft.

{ \red{personal funding sources}}
S.~Schmidt acknowledges support from DOE grant DE-SC0009999 and NSF/AURA grant N56981C.
AIM is advised by David W. Hogg and was supported by National Science Foundation grant AST-1517237.

In addition to packages cited in the text, analyses performed in this paper used the following software packages: \code{Numpy} and \code{Scipy} \citep{numpyscipy}, \code{Matplotlib} \citep{matplotlib}, \code{Seaborn} \citep{seaborn}, \code{minFunc} \citep{minfunc}, \code{pySkyNet} \citep{pyskynet}, and \code{photUtils} from the \lsst\ simulations package \citep{lsstphotutils}.

\input{desc-tex/ack/standard} % also available: key standard_short

\input{./tex/appendix}

\bibliography{main,lsstdesc}

\end{document}

% ======================================================================