### 2019-12-17 paper/contributions.tex

Author contributions are listed below. \\
S.J.~Schmidt: Co-led the project. (conceptualization, data curation, formal analysis, investigation, methodology, project administration, resources, software, supervision, visualization, writing -- original draft, writing -- review \& editing). \\
A.I.~Malz: Co-led the project, contributed to choice of metrics, implementation in code, and writing. (conceptualization, methodology, project administration, resources, software, visualization, writing -- original draft, writing -- review \& editing). \\
J.Y.H.~Soo: Ran ANNz2 and Delight, updated abstract, edited sections 1 through 6, added tables in Methods and Results, updated references.bib and added references throughout the paper. \\
I.A.~Almosallam: vetted the early versions of the data set and ran many photo-z codes on it, applied GPz to the final version and wrote the GPz subsection. \\
M.~Brescia: Main-ideator of MLPQNA and co-ideator of METAPHOR; modification of METAPHOR pipeline to fit the LSST data structure and requirements. \\
S.~Cavuoti: Co-ideator of METAPHOR, contributed to choice and test of metrics, ran METAPHOR, minor text editing. \\
J.~Cohen-Tanugi: contributed to running code, analysis discussion, and editing, reviewing the paper. \\
A.J.~Connolly: Developed the colour-matched nearest-neighbours photo-z code; participated in discussions of the analysis. \\
J.~DeRose: One of the primary developers of the Buzzard-highres simulated galaxy catalogue employed in the analysis. \\
P.E.~Freeman: Contributed to choice of CDE metrics and to implementation of FlexZBoost. \\
M.L.~Graham: Ran the colour-matched nearest-neighbours photo-z code on the Buzzard catalogue and wrote the relevant piece of Section 2; participated in discussions of the analysis. \\
K.G.~Iyer: assisted in writing metric functions used to evaluate codes. \\
M.J.~Jarvis: Contributed text on AGN to Discussion section and portions of GPz work. \\
J.B.~Kalmbach: Worked on preparing the figures for the paper. \\
E.~Kovacs: Ran simulations, discussed data format and properties for SEDs, dust, and ELG corrrections. \\
A.B.~Lee: Co-developed FlexZBoost and the CDE loss statistic, wrote text on the work, and supervised the development of FlexZBoost software packages. \\
G.~Longo: Scientific advise, test and validation of the modified METAPHOR pipeline, text of the METAPHOR section. \\
C.B.~Morrison: Managerial support; Discussions with authors regarding metrics and style; Some coding contribution to metric computation. \\
J.A.~Newman: Contributions to overall strategy, design of metrics, and supervision of work done by Rongpu Zhou. \\
E.~Nourbakhsh: Ran and optimized TPZ code and wrote a subsection of Section 2 for TPZ. \\
E.~Nuss: contributed to running code, analysis discussion, and editing,reviewing the paper. \\
T.~Pospisil: Co-developed FlexZBoost software and CDE loss calculation code. \\
H.~Tranin: contributed to providing SkyNet results and writing the relevant section. \\
R.H.~Wechsler: Project lead for Buzzard-highres simulated galaxy catalogue employed in analysis. \\
R.~Zhou: Optimized and ran EAZY and contributed to the draft. \\
R.~Izbicki: Co-developed FlexZBoost and the CDE loss statistic, and wrote software for FlexZBoost \\